Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-24 cav2017/cav2017.tex
Notations

Models \eqref{eq:ode} and \eqref{eq:plant} depend on external non-determinism in the form of input signals $u (t), u_k$, respectively. 
Feedback architectures can be employed to affect the properties and behaviours of the continuous process (the plant).   
%reduce the effect of parametric uncertainties and attenuate the effects of noise and disturbance. 
We are interested in the synthesis of digital feedback, 
as implemented on modern FPGAs or ICs. 
The most basic feedback architecture is the state feedback, 
where the control action $u_k$ (notice we work with the discretized signal) is computed by: 
%
\begin{equation}
\label{eq:controlaction}
u_k = r_{k} - K x_k. 
\end{equation}
%
Here $K \in \mathbb{R}^{m \times n}$ is a feedback gain matrix, 
and $r_{k}$ is a reference signal (again digital).   
%
The closed-loop model then takes the form 
\begin{align}
\label{eq:closedloopss}
x_{k+1} = ( A_d - B_d K ) x_k + B_d r_k.
\end{align}
The gain matrix $K$ can be set so that the closed-loop digitalized dynamics are shaped as desired, 
for instance according to a specific stability goal or around a specific dynamical behaviour \cite{astrom1997computer}. 
As argued later, in this work, we will target more complex objectives, such as quantitative safety requirements that are not typical in the digital control literature.

@@

In this study, we are not limited to the synthesis of digital stabilizing controllers -- 
a well known task in the literature on digital control system -- 
but target safety requirements with an overall approach that is sound and automated. 
More specifically, 
we require that the closed-loop system \eqref{eq:closedloopss} meets given safety specifications. 
A safety specification raises a requirement on the states of the model, 
such that the feedback controller (namely the choice of the gains in $K$) 
must ensure that the states never exist such requirement. 
Note that a stable, closed-loop system is not necessarily a safe system: 
indeed, the state values may leave the safe whilst they converge to the equilibrium, 
which is typical in the case of oscillatory dynamics. However, a safe system will always be stable.
In this work, the safety property represented by $\phi_{safety}$ is encoded as follows: 
%
\begin{equation}
\label{eq:safetyliteral}
\phi_{safety}\iff \forall k\geq 0, \bigwedge_{i=1}^{n}{\underline{x_{i}} \leq x_{i,k} \leq \overline{x_{i}}},
\end{equation}
%
%\addtodo{We consider $x_{i}^{-}$ and $x_{i}^{+}$ to be -1 and 1, respectively.}
where $\underline{x_{i}}$ and $\overline{x_{i}}$ are lower and upper bounds for the $i$-th coordinate $x_{i}$ of state $x\in \mathbb R^n$ at the $k$-th instant, respectively. 
This means that the states will always be within the $n$-dimensional hyper-box. 
%, which is typically defined by the safety specification.

@@

%%%-------------------------------
%%\subsection{Digital implementation of feedback controller}
%%\label{ssec:quantizationerror}
%%%-------------------------------
%%
%%\addtodo{HERE add a figure similar to Fig. \label{fig:observersystem} to explain digital features of controller.} 
%%
%%\addtodo{[The remainder of the section needs to be updated. ]}
%%The output and measurable of the plant is being sampled by an ADC, and since the full state of the
%%plant cannot be witnessed directly, the controller software includes an observer that 
%%estimates the state of the plant (see appendix for details on how such an observer
%%is constructed). In this work, we are assuming that all the states measurements are available for the state feedback purpose, and the state observer design may be disregarded.  
%%\addtodo{[Unclear what the remainder is for: ]}
%%We further presume that there are small tolerances in the system 
%%described in the following form
%%%
%%\begin{align*}
%%\mat{A} \in \mathcal{G}_A=\mat{S}\mat{J}\mat{S}^{-1} : \mat{J} \in \mathcal{G}_J
%%\end{align*}
%%\commentib{What is S?}
%%
%%Where $\mathcal{G}_J$ is a collection of eigenvalue ranges that define the frequency response of the system.
%%We presume that the internal model of the plant is an ARMA model, modified by a
%%state transform that has no tolerances. Tolerances in the eigenvectors will manifest themselves
%%as tolerances in the canonical forms outside of the row/column of interests and complicate the 
%%evaluation. Whilst it is possible to address these with this methodology, we will focus on the case
%%above for simplicity.

% \addtodo{HERE add a figure similar to Fig. \label{fig:observersystem} to explain digital features of controller.}

@@

\subsubsection{Effect on safety specification and stability}
Thus we can consider the effect of these errors on our safety and stability verification.

@@

The states are computed using the following equations. 
\begin{align*}
u_{k}&=-(\mathcal{F}_{\langle I_c,F_c \rangle}(K)\cdot\mathcal{F}_{\langle I_c,F_c \rangle}(x_{k})).  
% x_{k+1}&=A_d x_{k} + B_du_{k}
\end{align*}

This will induce two types of the error detailed above, first the truncation errors due to representing $x_k$ as ${\langle I_c,F_c \rangle}(x_{k})$, and second the rounding error introduce by multiplying $x_k$ and $K$. We represent these errors as non-deterministic noise.

A further error is caused by the representation of the plant dynamics.
\begin{align*}
\mathcal{F}_{\langle I_p,F_p \rangle}(u_{k})&=-(\mathcal{F}_{\langle I_c,F_c \rangle}(K)\cdot\mathcal{F}_{\langle I_c,F_c \rangle}(x_{k})).  \\
\mathcal{F}_{\langle I_p,F_p \rangle}(u_{k})(x_{k+1}) &=\mathcal{F}_{\langle I_p,F_p \rangle}(u_{k})(A_d) \mathcal{F}_{\langle I_p,F_p \rangle}(u_{k})(x_{k}) + \mathcal{F}_{\langle I_p,F_p \rangle}(u_{k})(B_d)\mathcal{F}_{\langle I_p,F_p \rangle}(u_{k})(u_{k})
\end{align*}

@@

At this point, we recap the the synthesis problem that we solve in this work:
find a digital controller $K$ (see Eq.~\ref{eq:controlaction})
that makes the closed-loop system safe for 
initial state $x_0$, reference signal $r_k$ and input $u_k$
as defined in Sec.~\ref{ssec:safety}. 
We consider non-deterministic initial states within a specified range
and the reference signal to be set to zero.  
When mapping back to the notation used for describing the general architecture 
of the program synthesizer, the controller $K$ denotes $P$ and 
$(x_0, u_k)$ represents $x$. 
While the controller's precision $\langle I_c,F_c\rangle$ is given, 
we can vary the plant's precision $\langle I_p,F_p\rangle$ such that 
$\mathbb{R}\langle I_p,F_p \rangle \supseteq \mathbb{R}\langle I_c,F_c \rangle$.

>>>>

Models \eqref{eq:ode} and \eqref{eq:plant} depend on external non-determinism in the form of input signals $u (t), u_k$, respectively. 
Feedback architectures can be employed to affect the properties and behaviours of the continuous process (the plant).   
%reduce the effect of parametric uncertainties and attenuate the effects of noise and disturbance. 
We are interested in the synthesis of digital feedback, 
as implemented on modern FPGAs or ICs. 
The most basic feedback architecture is the state feedback, 
where the control action $u_k$ (notice we work with the discretized signal) is computed by: 
%
\begin{equation}
\label{eq:controlaction}
u_k = r_{k} - K x_k. 
\end{equation}
%
Here $K \in \mathbb{R}^{m \times n}$ is a feedback gain matrix, 
and $r_{k}$ is a reference signal (again digital).   
%
The closed-loop model then takes the form 
\begin{align}
\label{eq:closedloopss}
x_{k+1} = ( A_d - B_d K ) x_k + B_d r_k.
\end{align}
The gain matrix $K$ can be set so that the closed-loop digitalized dynamics are shaped as desired, 
for instance according to a specific stability goal or around a specific dynamical behaviour \cite{astrom1997computer}. 
As argued later in this work, we will target more complex objectives, such as quantitative safety requirements that are not typical in the digital control literature.

@@

In this study, we are not limited to the synthesis of digital stabilizing controllers -- 
a well known task in the literature on digital control system -- 
but target safety requirements with an overall approach that is sound and automated. 
More specifically, 
we require that the closed-loop system \eqref{eq:closedloopss} meets given safety specifications. 
A safety specification raises a requirement on the states of the model, 
such that the feedback controller (namely the choice of the gains in $K$) 
must ensure that the states never exit such requirement. 
Note that a stable, closed-loop system is not necessarily a safe system: 
indeed, the state values may leave the safe space whilst they converge to the equilibrium, 
which is typical in the case of oscillatory dynamics. However, a safe system will always be stable.
In this work, the safety property represented by $\phi_{safety}$ is encoded as follows: 
%
\begin{equation}
\label{eq:safetyliteral}
\phi_{safety}\iff \forall k\geq 0, \bigwedge_{i=1}^{n}{\underline{x_{i}} \leq x_{i,k} \leq \overline{x_{i}}},
\end{equation}
%
%\addtodo{We consider $x_{i}^{-}$ and $x_{i}^{+}$ to be -1 and 1, respectively.}
where $\underline{x_{i}}$ and $\overline{x_{i}}$ are lower and upper bounds for the $i$-th coordinate $x_{i}$ of state $x\in \mathbb R^n$ at the $k$-th instant, respectively. 
This means that the states will always be within the $n$-dimensional hyper-box.

@@

@@

We will use $\mathcal{F}\langle I_c,F_c \rangle(x)$ to denote a real number $x$ represented in the fixed-point precision level of the controller, and $\mathcal{F}\langle I_p,F_p \rangle(x)$ to denote a real number $x$ represented in the fixed-point precision level of the plant model ($I_p \geq I_c \wedge F_p \geq F_c$). Thus any mathematical operations occurring in the digital controller will occur in the range of $\mathcal{F}\langle I_c,F_c \rangle$, and all other calculations will be carried out in the range of $\mathcal{F}\langle I_p,F_p \rangle$.

\subsubsection{Effect on safety specification and stability}
%Thus we can consider the effect of these errors on our safety and stability verification.

@@

The inputs are computed using the following equations. 
\begin{align*}
u_{k}&=-(\mathcal{F}_{\langle I_c,F_c \rangle}(K)\cdot\mathcal{F}_{\langle I_c,F_c \rangle}(x_{k})). 
\end{align*}

This will induce two types of the error detailed above, first the truncation errors due to representing $x_k$ as $\mathcal{F}{\langle I_c,F_c \rangle}(x_{k})$, and second the rounding error introduced by the multiplication. We represent these errors as non-deterministic noise.

A further error is caused by the representation of the plant dynamics.
\begin{align*}
x_{k+1} &=\mathcal{F}_{\langle I_p,F_p \rangle}(A_d) \mathcal{F}_{\langle I_p,F_p \rangle}(x_{k}) + \mathcal{F}_{\langle I_p,F_p \rangle}(B_d)\mathcal{F}_{\langle I_p,F_p \rangle}(u_{k})
\end{align*}

@@

At this point, we recap the the synthesis problem that we solve in this work:
find a digital controller $K$ (see Eq.~\ref{eq:controlaction})
that makes the closed-loop system safe for 
initial state $x_0$, reference signal $r_k$ and input $u_k$
as defined in Sec.~\ref{ssec:safety}. 
We consider non-deterministic initial states within a specified range
and the reference signal to be set to zero.  
When mapping back to the notation used for describing the general architecture 
of the program synthesizer, the controller $K$ denotes $P$ and 
$(x_0, u_k)$ represents $x$. 
While the controller's precision $\langle I_c,F_c\rangle$ is given, 
we can vary the plant's precision $\langle I_p,F_p\rangle$ such that 
$I_p \geq I_c \wedge F_p \geq F_c$.
